Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-09-08 Publications/2015-NeuroImage/responseLetter2.tex

\setlrmarginsandblock{3cm}{2.5cm}{*}
\setulmarginsandblock{2.5cm}{2.5cm}{*}
\checkandfixthelayout

\begin{document}

\section*{Response letter to the manuscript no. NIMG-15-1151}
\medskip
Original title: \emph{Active contours-driven registration method for the structure-informed segmentation of
  diffusion MR images}.

\noindent New title: \emph{{\color{blue} Surface}-driven registration method for the structure-informed segmentation of
  diffusion MR images}.

@@

\bigskip
\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-1151 submitted to NeuroImage.
We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
Please, find in the following our point-by-point answer to the concerns raised by the reviewers,
  along with the description of the corresponding changes in the manuscript.

\bigskip
\bigskip
\subsection*{Reviewer \#1:}
\begin{reviewpoint}
The paper describes an approach for the delineation of selected anatomical structures ([...]) in multi-shell diffusion MR images
  by registering a segmented T1 weighted reference image with fractional anisotropy and apparent diffusion coefficient maps derived
  from the diffusion weighted images.

The approach corresponds fully to the usual registration of an atlas to the image to be segmented, leaving me rather clueless why do
  they call the method ``simultaneous registration and segmentation''.

\end{reviewpoint}
\reply{
The first sentence of the comment is a rather accurate interpretation of our work.
However, it is important to clarify that our method maps anatomically-correct surfaces of the same subject into dMRI space,
  thereby segmenting this target space.
The method works regardless how the surfaces were extracted, or if they were extracted from T2w images or manually delineated.
In this work, the surfaces are extracted from T1w for two reasons: 1) it is the widely-accepted MR sequence for such extraction,
  and 2) the structural information used in reconstruction, tractography and network definition of connectivity extraction pipelines
  is usually defined in T1w-space as well.
We understand that the manuscript was not clear to this respect, and therefore included the following changes:
\begin{enumerate}
\item Changed abstract, more oriented on the problem now.
\item Re-structured introduction: lines 1-14 state the background and the problem more clearly, lines 15-27 review
  the literature existing about direct segmentation of dMRI data and lines 28-55 describe the alternative approach
  by registering structural images and the problem of EPI distortions.
  The methodological background of active-contours and registration-segmentation methods has been relocated in
    the beginning of section 2, lines 73-85.
\end{enumerate}

Our method can be seen as a segmentation framework using active contours without edges in which the evolution
  of the contours is governed by an elastic registration method;
or as a nonlinear registration method which cost function is built on top of a segmentation model of active contours
  without edges.
In our opinion, the presented approach is mainly a segmentation method.
However, since the surfaces are supposed to be very accurate and defined in a different coordinate system,
  the method implicitly registers that space to the target image.
Some comments of Reviewer \#2 are very similar to this one (particularly \autoref{com:11}).
}

\begin{reviewpoint}
The paper describes an approach [...] (quite some are listed as possible examples but results are only reported for the white matter,
  deep gray matter and the pial surface) in multi-shell [...].
\end{reviewpoint}
\reply{
We have defined clearly that the choice of surfaces to segment the target data is crucial, and define the
  segmentation model that drives the registration process (lines 110-113 are new).
We have also improved former Figure 4, and brought it earlier in the paper (now it is Figure 2).
Now, the figure shows how the segmentation model works: A is the ideal segmentation, B shows the model in
  the initialization settings, and C shows the model after using \emph{regseg}, demonstrating how the covariance
  matrices for each region are reduced.
Since the former visualization was a bit clumsy, we have extracted the distribution of diffusion features by region on
  the right side of each row.
}

\begin{reviewpoint}
The proposed algorithm is rather standard, relying on a region-based elastic contour deformation approach.
The cost function is derived in the usual Bayesian framework, relying on a multivariate normal distribution prior for the image
  intensities and a Tikhonov regularizer enforcing a predefined normal distribution over the distortions along the imaging axes.
In this respect the novelty of the paper can be questioned, while the application context and the very promising results may still
  justify a publication in NeuroImage.
\end{reviewpoint}
\reply{
We agree with the reviewer on that the novelty of the methodological elements of our approach is limited.
However, to our knowledge there exists no alternative method with a paired combination of the following elements:
\begin{itemize}
\item Region based (i.e. without edges) active contours model
\item The use of surfaces as prior knowledge to initialize the segmentation model,
  with property of allowing precise registration if the surface is accurately representative
  of structures being aligned.
  Moreover, the multi-phase definition of levelsets is dual to our surface-nesting model.
\item The anisotropic regularization model of choice, since most of the levelset frameworks
  use Gaussian smoothing of the deformation field.
\item Even though the shape-gradients theory was proposed back in the early 2000's, atlas-based
  registration, and registration-segmentation methods have been generally implemented on levelsets.
\end{itemize}
To our knowledge, the application of a framework as such is new and an important contribution is done
  with the proposed segmentation model of the FA and ADC maps.
We would have appreciated if the reviewers would have provided specific references being close to our approach
  beyond those given in lines 43-57 of the original paper, so that we could point to the differences of our
  method.
For the sake of clarity, we have removed the methodological background from introduction to section 2, lines 73-85.
}

\begin{reviewpoint}
However, in such a case the paper should be thoroughly revised to address the following major issues:
1. The assumption of independence between pixels is quite usual but obviously wrong.
The authors may want to comment on the consequences.
\end{reviewpoint}
\reply{
The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
We have supported the assumptions with new lines 105-107:
\textit{Even though the features being segmented are not generally i.i.d., the spatial interdependency of voxels is
implicitly supported by the piecewise smooth partition of the space $\boldsymbol\Omega$.
In Figure 2 it is shown that marginal distributions of data can be approximated by multivariate normal distributions.}
}

\begin{reviewpoint}
2. It is unclear, under which conditions enforcing the optimality condition guarantees that the region identified corresponds
  to the targeted one (it probably strongly depends on its intensity properties and those of neighbouring structures).
A corresponding analysis of the goal function and the presence of (undesired) local minima is missing.
Figure S1 is not particularly helpful in this respect.
Some information about the initialization of the distortion map should also be given.
\end{reviewpoint}
\reply{
We have improved former Figure 4 (now, Figure 2) to clarify that the model strongly depends on the choice of surfaces.
Adding more surfaces makes the model more robust to local minima at the cost of largely increased computation times.
We agree with the reviewer on that Figure S1 is not illustrative of the metric for free-deformation fields, but
  our intention was to show the performance of the method in realistic settings.
This implies distortions with nonzero components on the phase-encoding axis (the $y$-axis in our experiments), and
  therefore Figure S1 becomes representative since it samples the distortions space covering plausible deformations
  for both encoding directions (Anterior-to-posterior and posterior-to-anterior).
We have included information about the maximum distortion for which the metric was sampled.
}

@@

\begin{reviewpoint}
3. I miss a reasonable description of the parameter settings necessary for generating the results presented.
It is not even clear if all experiments have been performed by the same parameter settings.
A reasonable sensitivity analysis is also missing.
\end{reviewpoint}
\reply{
We have made a strong effort in providing reproducible experiments using openly available data.
Therefore, parameter settings are publicly available in GitHub.
Particularly, under folder \url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data}
  there are several configuration files.
Some parameters are also discussed in the Supplemental Materials, Section S2.
However, we understand that a better description of the parameter settings should be included within
  the manuscript and modified it accordingly as follows: {\color{red} say what here}.
}

\begin{reviewpoint}
4. I could not understand how the ground truth can be generated for the real datasets.
As far as I can see, in these cases there is a (probably unknown) distortion between the T1 weighted and the
  diffusion weighted images.
How an emulated distortion can get around this fact?
\end{reviewpoint}
\reply{
As commented in line 160, we are using some ``minimally preprocessed'' images from the database of the
  Human Connectome Project (HCP).
Particularly, the dMRI datasets from these preprocessed images are given corrected for distortions
  and registered in structural space.
Therefore, there could be a residual misregistration between the T1w and the dMRI data, but it should be
  very minimal.
We have modified Figure 3 to reflect this aspect.
}

\begin{reviewpoint}
5. The vertex-wise error distributions are not only skewed, but quite clearly structured im many cases
  (especially for the phantom images).
What is the underlying reason?
\end{reviewpoint}
\reply{
We have added the following in the Discussion section (lines 262-267):

\textit{The distributions of errors along surfaces vary importantly depending on the shape of the
  phantom (see Figure 5B).
The misregistration error of the ``gyrus'' phantom showed a much lower spread than that for the other shapes.
We argue that the symmetry of those other shapes posed problem in driving the contours towards the appropriate
  region and producing some ``sliding'' effect between the faces of the surface and their ground-truth position.
This effect should not be present in real datasets, thanks to the very convoluted cortical layer, and the
  directional restriction of distortion.}
}

\bigskip
\bigskip
\subsection*{Reviewer \#2:}
\begin{reviewpoint}
This paper seems to present a method that can perform wM-GM segmentation and simultaneously estimate a deformation field to
  correct for EPI distortion.
Apparently, the idea is to use nested WM-GM surfaces, as well as other surfaces, to provide better cortical registration than
  a competing method, which is not clearly described.
The overall goal and motivation of the paper are not clearly explained, and almost all related references for EPI distortion
  correction are missing.
It is unclear if the deformation field solution is restricted to the phase encode direction.
It is possible that this is a very useful method, as the results seem to indicate, but the paper is unfortunately too unclear to
  make a full judgement about the method.
\end{reviewpoint}
\reply{
We thank the reviewer for pointing out that the overall goal is not clearly explained.
We have replaced the term ``active-contours'' by ``surfaces'' whenever possible.
We have re-organized the introduction fully, as suggested (see reply to \autoref{com:1} for a brief explanation).
We have included references of the main EPI distortion corrections (see lines 31-43).

The \emph{regseg} tool is designed to allow any displacement, as demonstrated with the phantoms.
However, as the reviewer points out, the deformation field is completelly restricted to the phase-encode direction
  for the real datasets.
We have added lines 204-206 to clarify this point.
}

\begin{reviewpoint}
abstract: Not immediately clear what this means: ``active contours without edges extracted from structural images''
  (Where do the edges come from? What modalities are actually input to regseg?)
\end{reviewpoint}
\reply{Active contours without edges \citep{chan_active_2001} is a framework for subpixel segmentation of images in
  which the contours evolve enclosing homogeneous regions inside.
Region-based active contours is an alternative terminology.
As previously described in \autoref{com:1} and \autoref{com:2}, we have made an effort to clearly state that the outputs
  to our method are a set of nested surfaces and a target multivariate image, that for the case of dMRI data comprehends
  FA and ADC.
The manuscript has been re-phrased accordingly in the title, abstract and introduction.}

@@

\begin{reviewpoint}
I note that the text in the abstract seems to have no correspondence with the graphical abstract.
The text abstract mentions nothing about diffusion MRI, FA, ADC, or phase maps, so it is unclear how these may be used in the method.
These images figure prominently in the graphical abstract, which may be a better overview than the text abstract.
\end{reviewpoint}
\reply{We agree with the reviewer, we have updated the text abstract, and proposed a completely new graphical abstract.}

\begin{reviewpoint}
Intro line 6-7. I am sure you mean to say that the resolution of dMRI is much LOWER than the images microstructure (not higher).
  larger voxels == lower resolution
\end{reviewpoint}
\reply{Absolutely, we have fixed the manuscript accordingly.}

\begin{reviewpoint}
Intro 9. Orientations is the usual term for diffusion weighting, rather than ``angles''.
\end{reviewpoint}
\reply{We agreed, and updated the manuscript accordingly.}

\begin{reviewpoint}
``These limitations prevent segmentation in the native dMRI space''  I disagree.
In fact, many methods have been used to segment fiber tracts, then to quantitatively measure them,
  in the dMRI space.
It is true the correspondence with T1 is imperfect, but segmentation of fiber tracts is possible.
\end{reviewpoint}
\reply{
The reviewer is right, we have changed the paragraph about dMRI segmentation (lines 15-27) to
  be more precise.
We agree with the reviewer that there actually exist many methods to segment fiber tracts.
However, our method is intended to support fiber tracking methods requiring detailed anatomical
  information, such as \citep{smith_anatomicallyconstrained_2012}.
We have supported the applications more clearly on lines 11-14.
Clustering tractograms could be another potential application of \emph{regseg},
  but we agree with the reviewer that current methods perform well without
  additional structural information.
}

@@

\begin{reviewpoint}
Please define what you mean here by segmentation.
It seems there is an assumption that it means specifically segmentation via registration, where the
  structures are defined using T1 images, but that is not stated anywhere.

Now on line 19 it seems the goal of the segmentation may simply be the segmentation of the WM.
So far, I have no idea what the overall goal of the approach is.
\end{reviewpoint}
\reply{
This comment is very related to \autoref{com:14}.
We have modified the abstract and introduction accordingly.
}

\begin{reviewpoint}
Line 24. clustering of what?
\end{reviewpoint}
\reply{
This comment is a consequence of the two previous comments.
The reviewer is right in the sense that we did not clearly state what we are segmenting, and therefore
  the references in line 24 are a bit orphaned.
All methods reviewed in lines 19-28 cluster/segment/classify \emph{voxels} in dMRI space based on
  one or another feature derived from dMRI data.
We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15}, and specifically
  rewritten lines 19-28 to explicitly state what those methods are clustering.
}

@@

\begin{reviewpoint}
Now in the paragraph beginning on line 29, it seems the segmentation of WM/GM interface is of interest.
Still, I have no idea if this is the goal of the approach, or why.

Line 39. I note that there are many approaches, in addition to nonlinear T2->b0, that have been proposed
  for correction of EPI distortions in dMRI.
If that is the goal of this approach, the entire intro should be rewritten focusing on this related work.
However, I do not know yet what the goal is.

The entire Introduction should be rewritten to improve organization and clearly state the paper's goal much
  earlier (in abstract and introduction).
Truly related work should be described in more detail, and all of the unrelated references should be removed.

Line 58. It seems the paper goal may be stated here. It is still completely unclear what type of segmentation should be achieved.
\end{reviewpoint}
\reply{
Again, the reviewer finds that the main goal of the paper is not clearly stated.
We understand that modifications derived from the previous comments satisfy the answer to the concern,
  and changes include abstract and introduction.
We have stated more clearly the importance of the set of surfaces being registered to produce a reliable
  segmentation model of the two diffusion features selected: FA and ADC
}

@@

\begin{reviewpoint}
It seems by line 69 that EPI distortion correction and segmentation by registration are to be performed jointly
  in some fashion.
There are many, many references for distortion correction that are missing from this introduction and that must
  be added.
See for example multiple papers by Carlo Pierpaoli's group, including this most recent:

Irfanoglu, M. Okan, et al. ``DR-BUDDI (Diffeomorphic Registration for Blip-Up blip-Down Diffusion Imaging) method for correcting echo planar imaging distortions.''
  NeuroImage 106 (2015): 284-299.
\end{reviewpoint}
\reply{
We have added the suggested reference (among other of interest) in lines 204-206.
In lines 59-61 we have stated the differences of \emph{regseg} w.r.t. EPI distortion corrections.
}

@@

\begin{reviewpoint}
The majority of the EPI distortion correction methods use T2 and B0 due to similar contrast. Why was T1 chosen in this framework?

Suddenly in Section 2 it becomes clear that there are surfaces being registered.
Perhaps this should have been obvious from the title, which mentioned active contours, but before line 72 it was not clearly
  stated that the method was posed as a surface registration problem.
\end{reviewpoint}
\reply{
We have updated the title to represent this aspect.
We have also included all previous changes to fulfill this request.
}

@@

\begin{reviewpoint}
Was the warp restricted to the phase encode direction?
If so, the r in eq (1) should be restricted to this direction.
I do not see this mentioned anywhere near the equations in the Methods section.
If the solution found for the deformation field was not restricted to the phase encode direction,
  then this is not a reasonable way to perform EPI distortion correction. Please clarify.

[...]

I still see nothing in the results indicating that the estimated distortions were restricted to the phase encode direction.
If not, the added degrees of freedom might result in improved surface segmentation, but the distortion field would not correspond at
  all to the physical problem under study.
Furthermore, the dMRI gradients would have to be re-oriented independently at each voxel, and this is not currently possible in any
  software I know of.
\end{reviewpoint}
\reply{
The displacements field supporting the deformation model is not restricted to the phase encoding direction by default.
Therefore, equations in Methods section are presented under such conditions.
The proof of concept presented in the first experiment with phantoms are conducted without preferential directions
  for the deformation.
However, the point raised about the distortions derived from susceptibility inhomogeneity in dMRI data is also true.
Therefore, experiments with real data mimicking such distortions are performed with the deformation aligned with the phase encoding
  axis.
New lines 203-205 now state this point clearly.
}

\begin{reviewpoint}
The experiments seem reasonable and clearly presented in general. But far too little detail is given on the B0-T2 comparison method.
\end{reviewpoint}
\reply{
We thank the reviewer for positively weighing our experiments and results.
As stated in line 214, we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
A footnote in line 215 directly links the settings for this method.
}

\begin{reviewpoint}
Please define the phrase ``active contours without edges''.
\end{reviewpoint}
\reply{We hope this point is answered in \autoref{com:10}, and tryied to avoid this concept
  coming from the 2D processing field.
Now, we refer to the contours as surfaces whenever possible.
Regarding the ``without edges'' feature, we have clarified how the surfaces enclose homogeneous
  regions (and therefore, they are not used to search for image edges).
}

### 2015-09-08 Publications/2015-NeuroImage/02-abstract.tex
v1.0 of resubmission

% -*- root: 00-main.tex -*-
Current applications of whole-brain tractography to \acrlong*{dmri} data require highly precise
  delineations of anatomical structures, which are usually projected from \acrlong*{t1} images
  by registration.
In this study, we propose \regseg{}, which is a simultaneous segmentation and registration method
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-form deformation field supported by the B-spline basis
  to optimally map the contours onto the data in the target space.
We tested the functionality of \regseg{} using four digital phantoms warped with known and
  randomly generated deformations, where subvoxel accuracy was achieved.
We then applied \regseg{} to a registration/segmentation task using 16 real diffusion MRI
  datasets from the \acrlong*{hcp}, which were warped by realistic and nonlinear distortions that are typically
  present in these data.
We computed the misregistration error of the contours estimated by \regseg{} with respect
  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
We also compared the performance of our proposed method with a widely used registration tool, which showed
  that \regseg{} outperformed this method in our settings.

>>>>

% -*- root: 00-main.tex -*-
Current methods using diffusion MRI to map the microstructure and connectivity of the
  human brain increasingly require precise delineations of anatomical structures.
The problem is typically solved by projecting the structural information extracted
  in T1-weighted images, by means of nonlinear registration since diffusion
  MRI data present geometrical distortions.
Here we propose \regseg{}, which segments homogeneous regions within multivariate
  images by projecting a set of nested surfaces extracted from structural images.
The surfaces evolve through a free-form deformation field supported by the B-spline basis
  to optimally map the contours onto the data in the target space.
We tested the functionality of \regseg{} using four digital phantoms warped with known and
  randomly generated deformations, where subvoxel accuracy was achieved.
We proposed a set of surfaces to segment the fractional anisotropy and the apparent diffusion
  coefficient maps derived from diffusion MRI datasets.
These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real
  diffusion MRI datasets from the Human Connectome Project that were warped by realistic and nonlinear
  distortions.
We computed the misregistration error of the surfaces estimated by \regseg{} with respect
  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
We also compared the performance of the proposed method with a widely used registration tool, which showed
  that \regseg{} outperformed this method in our settings.

### 2015-09-08 Publications/2015-NeuroImage/03-introduction.tex
v1.0 of resubmission

The tool \emph{bbregister} \citep{greve_accurate_2009} has become a rather standard linear
  registration method to be used along with \gls*{epi} corrections.
It uses 3D active contours with edges\footnote{A three-dimensional active contour is a surface
  and it is ``with edges'' when the contours evolve in search of abrupt intensity steps (the edges) within
  the target image.} to search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
The \lowb{} image only includes a detectable frontier for the pial surface, and thus
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
  regions that are typically warped from the boundary search.
However, because the distortion is not considered, it must be addressed separately.
The nonlinear registration approach has been successfully solved with \gls*{t2b} methods,
  which align the \gls*{t2} and \lowb{} images by exploiting their similarity
  \citep{kybic_unwarping_2000,studholme_accurate_2000}.
However, further rigid registration of the \gls*{t1} and \gls*{t2} images is still required to
  map the anatomical information onto \gls*{dmri}-space.

Here we show a method that registers anatomically-correct surfaces to multivariate images,
  thereby implicitly segmenting the \gls*{dmri}-space into those structures enclosed by the contours
  of the subject's brain.
% It is important that surfaces are defined in the coordinate system of the image used as
%   source of structural information (typically, but not restricted to, \gls*{t1} images), otherwise segmentation performs
%   correctly but the mapping found in registration does not represent any valuable information.
The hypothesis tested in the present study is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, effectively establishing the
  necessary mapping to project structural information from anatomical images, increasing the
  geometrical accuracy of the overall process, and avoiding data resampling of \gls*{epi}
  correction methods.
Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges
  \citep{chan_active_2001}, which evolve over a free-form deformation field of B-spline
  basis functions.
The surfaces are nested and they enclose regions with homogeneous diffusion features
  (the \gls*{fa} and \gls*{adc} maps for the case of \gls*{dmri} data).
Optimization is performed using a descent strategy with explicit shape gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, unlike most previous methods, \regseg{} does not implement level sets.
Since the \gls*{epi} distortion is aligned with the \gls*{pe} axis (see
  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer for
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluate \regseg{} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
  based on \gls*{dmri} data.
The evaluation includes a comparison of \regseg{} to an in-house implementation of
  the \gls*{t2b} correction.

>>>>

The tool \emph{bbregister} \citep{greve_accurate_2009} has been widely adopted as the standard
  registration method to be used along with the \gls*{epi} correction of choice.
It implements a linear mapping, and uses 3D active contours with edges\footnote{A three-dimensional active contour
  is a surface and it is ``with edges'' when the contours evolve in search of abrupt intensity steps
  (the edges) within the target image.} to search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
The \lowb{} image only includes a detectable frontier for the pial surface, and thus
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
  regions that are typically warped from the boundary search.
However, because the distortion is not considered, it must be addressed separately.
The nonlinear registration approach has been successfully solved with \gls*{t2b} methods,
  which align the \gls*{t2} and \lowb{} images by exploiting their similarity
  \citep{kybic_unwarping_2000,studholme_accurate_2000}.
However, further rigid registration of the \gls*{t1} and \gls*{t2} images is still required to
  map the anatomical information onto \gls*{dmri}-space.

Here we show a method that registers anatomically-correct surfaces to multivariate images,
  thereby implicitly segmenting the \gls*{dmri}-space into those structures enclosed by the contours
  of the subject's brain.
% It is important that surfaces are defined in the coordinate system of the image used as
%   source of structural information (typically, but not restricted to, \gls*{t1} images), otherwise segmentation performs
%   correctly but the mapping found in registration does not represent any valuable information.
The hypothesis underlying is that the registration and segmentation problems in \gls*{dmri}
  can be solved simultaneously with three key benefits:
  1) effectively establishing the necessary mapping to project structural information from anatomical images,
  2) increasing the geometrical accuracy of the overall process,
  and 3) avoiding data resampling of \gls*{epi} correction methods.
Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges
  \citep{chan_active_2001}, which evolve over a free-form deformation field of B-spline
  basis functions.
The surfaces are nested and they enclose regions with homogeneous diffusion features
  (the \gls*{fa} and \gls*{adc} maps for the case of \gls*{dmri} data).
Optimization is performed using a descent strategy with explicit shape gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, unlike most previous methods, \regseg{} does not implement level sets.
Since the \gls*{epi} distortion is aligned with the \gls*{pe} axis (see
  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer for
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluate \regseg{} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
  based on \gls*{dmri} data.
The evaluation includes a comparison of \regseg{} to an in-house implementation of
  the \gls*{t2b} correction.
